The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2023-07-03 analysis/paper/paper.qmd
Create word embeddings functions

Describe word embeddings ...

```{r word-embeddings}
#| include: false

# Prepare tidy data frame for word embeddings
tidy_emu_theses <- emu_theses |> 
  select(title, text_clean) |> 
  unnest_tokens(word, text_clean) |> 
  add_count(word) |> 
  filter(n >= 50) |> 
  select(-n)

# Create a nested data frame with one row per thesis
nested_words <- tidy_emu_theses |> 
  nest(words = c(word))

# Create sliding window function to calculate skipgram probabilities,
# that is, how often each word occurs near each other word.
# A small window size of 3-4 shows how a word is used, while a larger  
# window size of, say, 10 gives more information about the domain topic.
slide_windows <- function(tbl, window_size) {
  # Iterate through an input vector with a fixed window size
  skipgrams <- slider::slide(
    tbl,
    ~.x,
    .after = window_size - 1,
    .step = 1L,
    .complete = TRUE  # Only evaluate function on complete windows
  )

  # Run even when unsuccessful
  safe_mutate <- safely(mutate)

  # Map safe_mutate to skipgrams 
  out <- map2(skipgrams,
              1:length(skipgrams),
              ~ safe_mutate(.x, window_id = .y))

  out |> 
    transpose() |>  # Transpose `result` and `error` into columns,
    pluck("result") |>  # only keep `result`,
    compact() |>  # remove any NULL values [is this for the last windows?],
    bind_rows()  # and bind all rows into a single data frame.
}
```

# To reduce computation time, we use the {furrr} 
# and underlying {future} packages for parallel processing
future::plan(multisession)

# To calculate how often words occur on their own or tigether with other words
# we use point-wise mutual information (PMI)
tidy_pmi <- nested_words |> 
  mutate(words = future_map(words, slide_windows, 4L)) |> 
  unnest(words) |> 
  unite(window_id, title, window_id) |> 
  pairwise_pmi(word, window_id)

@@

@@

# Determine word vectors using SVD, a method for dimensionality reduction
# The resulting matrix will be of size n_voc * n_dim instead of n_voc * n_voc
tidy_word_vectors <- tidy_pmi |> 
  widely_svd(
    item1, item2, pmi,
    nv = 100, maxit = 1000
  )

# Calculate nearest neighbours of a given word
nearest_neighbors <- function(df, token) {
  df |> 
    widely(
      ~ {
        y <- .[rep(token, nrow(.)), ]
        res <- rowSums(. * y) /
          (sqrt(rowSums(. ^ 2)) * sqrt(sum(.[token, ] ^ 2)))

        matrix(res, ncol = 1, dimnames = list(x = names(res)))
      },
      sort = TRUE
    )(item1, dimension, value) |> 
    select(-item2)
}

tidy_word_vectors |> 
  nearest_neighbors("price")
```

# The first n components with top words found in the corpus of theses
tidy_word_vectors |> 
  filter(dimension <= 24) |> 
  group_by(dimension) |> 
  top_n(12, abs(value)) |> 
  ungroup() |> 
  mutate(item1 = reorder_within(item1, value, dimension)) |> 
  ggplot(aes(item1, value, fill = dimension)) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~dimension, scales = "free_y", ncol = 4) +
  scale_x_reordered() +
  coord_flip() +
  labs(
    x = NULL,
    y = "Value",
    title = "First 24 principal components for text of theses",
    subtitle = "Top words contributing to the components that explain the most variation"
  )

>>>>

### Word embeddings

[Describe word embeddings ...]

```{r word-embeddings-pmi}
#| include: false

# To what extent do words co-occur with other words?
tidy_pmi <- get_pmi(emu_theses, title_col = "title", text_col = "text_clean")

@@

```{r word-embeddings-svd}

# Determine word vectors using SVD, a method for dimensionality reduction
tidy_word_vectors <- tidy_pmi |> 
  widely_svd(
    item1, item2, pmi,
    nv = 100, maxit = 1000
  )

[What terms should be checked with the function below?]

```{r word-embeddings-nn}

# Calculate nearest neighbours of a given word
tidy_word_vectors |> 
  nearest_neighbors("resilience")

The first [?] components... [what do they show?]

@@

# The first n components with top words found in the corpus of theses
visualise_first_n_components(tidy_word_vectors, 24, 12)
